Preprocessing and exploration of the narwhal data for the analysis with smoothSDE

### 2024-04-05 R/preprocessing/data_preprocessing.Rmd
Preprocessing updates

```{r import packages,message=FALSE}

@@

```{r read raw data, cache=TRUE}
# Set the path to the directory containing the data
par_dir=dirname(dirname(getwd())) #parent directory 
narwhal_data_path <- file.path(par_dir,"Data", "Narwhals")

@@

```{r preprocess data 1, cache=TRUE}

@@

```{r plot trials, cache=TRUE}

@@

```{r define ExpShipI and ExpShipT, cache=TRUE}

@@

```{r ggplot ship exposure, cache=TRUE}
#plot exposure
pExpAll=ggplot(data,aes(time,ExpShip,col=ID))+geom_line()+
  scale_color_viridis_d(labels=c("A1","A2","A3","A4","A5","A6"))

@@

```{r plot exposure per individual,cache=TRUE}

@@

```{r preprocess data 2, cache=TRUE}

@@

```{r}
times_histo=ggplot(dtimes, aes(x =60*delta,col=ID)) +
  geom_histogram(fill="white",bins=100)+facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  scale_color_viridis_d(guide="none")+
  theme(legend.position="top")+xlab("Time step (min) ")

@@

```{r milne land map, cache=TRUE}
height <- max(dataprep$Lat) - min(dataprep$Lat)
width <- max(dataprep$Long) - min(dataprep$Long)
borders <- c(bottom  = min(dataprep$Lat), 
                 top     = max(dataprep$Lat),
                 left    = min(dataprep$Long),
                 right   = max(dataprep$Long))

@@

```{r trajectories map, cache=TRUE}

@@

```{r extract coastline and land,cache=TRUE}

@@

#get the coastline geometry from the geojson file
coastline_utm<-st_read(file.path(greenland_data_path,"filtered_coastline_utm.geojson"))
st_crs(coastline_utm)="+init=EPSG:32626"

#get land polygons
land_utm<-st_read(file.path(greenland_data_path,"filtered_land_utm.shp"))

@@

```{r compute nearest shore points,cache=TRUE}

@@

```{r compare distances to shore}
plot(res_land$min_distances/1000-dataprep$DistanceShore)
plot(res_coastline$min_distances/1000-dataprep$DistanceShore)
```
```{r summary distances to shore}
summary(res_land$min_distances)
summary(res_coastline$min_distances)
summary(dataprep$DistanceShore*1000)
```
Coastline distances are a bit closer to original distances in the dataset.
Anyway, we shall keep distance computed from the land in the sequel, since the land polygons will be the reference domain
for constrained movement of the narwhals.

@@

df=data.frame(DistanceShore=res_land$min_distances[res_land$min_distances<100 & res_land$min_distances>0  ])
histo=ggplot(df, aes(x =DistanceShore)) +
  geom_histogram(fill="cyan",bins=100,color="black")+xlab("Computed distance to shore (m)")
```

```{r}
histo
```

@@

```{r empirical velocities and deviation angles, cache=TRUE}

>>>>

```{r libraries,message=FALSE}

@@

```{r read raw data}
# Set the path to the directory containing the data
par_dir=dirname(dirname(getwd())) #parent directory 
narwhal_data_path <- file.path(par_dir,"Data", "Narwhals")

@@

```{r preprocess data 1}

@@

```{r plot trials}

@@

```{r define ExpShipI and ExpShipT}

@@

```{r ggplot ship exposure}
#plot exposure
pExpAll=ggplot(data,aes(time,ExpShip,col=ID))+geom_line()+
  scale_color_viridis_d(labels=c("A1","A2","A3","A4","A5","A6"))

@@

```{r plot exposure per individual}

@@

```{r preprocess data 2}

@@

```{r time steps histogram}
times_histo=ggplot(dtimes, aes(x =60*delta,col=ID)) +
  geom_histogram(fill="white",bins=100)+facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  scale_color_viridis_d(guide="none")+
  theme(legend.position="top")+xlab("Time step (min) ")

@@

```{r milne land map}
height <- max(dataprep$Lat) - min(dataprep$Lat)
width <- max(dataprep$Long) - min(dataprep$Long)
borders <- c(bottom  = min(dataprep$Lat), 
                 top     = max(dataprep$Lat),
                 left    = min(dataprep$Long),
                 right   = max(dataprep$Long))

@@

```{r trajectories map}

@@

```{r extract coastline and land}

@@

#get the coastline geometry from the geojson file
coastline_utm<-st_read(file.path(greenland_data_path,"coastline_utm.geojson"))
st_crs(coastline_utm)="+init=EPSG:32626"

#get land polygons
land_utm<-st_read(file.path(greenland_data_path,"land_utm.shp"))

@@

```{r compute nearest shore points}

@@

```{r compare distances to shore}
plot(res_land$min_distances/1000-dataprep$DistanceShore)
plot(res_coastline$min_distances/1000-dataprep$DistanceShore)
```

```{r summary distances to shore}
summary(res_land$min_distances)
summary(res_coastline$min_distances)
summary(dataprep$DistanceShore*1000)
```
Coastline distances are a bit closer to original distances in the dataset.
Anyway, we shall keep distance computed from the land in the sequel, since the land polygons will be the reference domain
for constrained movement of the narwhals.

@@

```{r distribution of points close to the shore}

df=data.frame(DistanceShore=res_land$min_distances[res_land$min_distances<100 & res_land$min_distances>0  ])
histo=ggplot(df, aes(x =DistanceShore)) +
  geom_histogram(fill="cyan",bins=100,color="black")+xlab("Computed distance to shore (m)")

histo
```

@@

```{r empirical velocities and deviation angles}
